---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    fig_caption: yes
    keep_md: yes
---
## Requirements
Please make sure you have dplyr, ggplot2 and lubridate package loaded
```{r,}
library(dplyr)
library(ggplot2)
library(lubridate)
```

## Loading and preprocessing the data
Loading the file and reformat date column to date format
```{r}
activity <- read.csv("./Data//activity.csv")
activity <- mutate(activity,date = as.POSIXct(date, format="%Y-%m-%d"))
```

## What is mean total number of steps taken per day?

1. Create a grouping by date
2. Summarize the steps
3. Draw histogram of steps by day
```{r}
activity_group <- group_by(activity,date)
sumStepsByDay <- summarize(activity_group, sumSteps = sum(steps, na.rm = TRUE))
qplot(sumSteps, data = sumStepsByDay, xlab = "Sum of Steps", binwidth = 450)
```

Show Summary statistics
```{r}
summary(sumStepsByDay$sumSteps)
```

## What is the average daily activity pattern?
Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

1. Create a grouping by interval
2. Creating the mean by interval accross all days
3. Draw time series plot
```{r}
activityIntervalGroup <- group_by(activity,interval)
meanStepsbyInterval <- summarize(activityIntervalGroup, meanSteps = mean(steps, na.rm = TRUE))
plot(meanStepsbyInterval$interval,meanStepsbyInterval$meanSteps,type="l", xlab = "5-minute interval", 
        ylab="Average number of steps")
```

Showing the intervals with the maximum number of steps
```{r}
arrange(meanStepsbyInterval,desc(meanSteps))
```

## Imputing missing values
Show summary and Missing Values (NA's)
```{r}
summary(activity)
```


A new dataset tidyActivity is created. 
Strategy for imputing missing values: NA values are replaced with the mean of the steps by interval.
```{r}
tidyActivity <- group_by(activity,interval) %>% mutate(runningmean=mean(steps,na.rm=TRUE)) %>% mutate(steps=ifelse(is.na(steps),runningmean,steps)) %>%
        select(steps,date,interval)
```

Create histogram for tidydataset:

1. Create a grouping by date
2. Summarize the steps
3. Draw histogram of steps by day
```{r}
tidyActivityGroup <- group_by(tidyActivity,date)
tidyActivitySumStepsGroup <- summarize(tidyActivityGroup, sumSteps = sum(steps, na.rm = TRUE))
qplot(sumSteps, data = tidyActivitySumStepsGroup, xlab = "Sum of Steps", binwidth = 450)
```

Show Summary statistics
```{r}
summary(tidyActivitySumStepsGroup$sumSteps)
```

###Conclusion
*Do these values differ from the estimates from the first part of the assignment?*

Yes, the minimum, 1.st Qu., Median, Mean and 3rd Qu are different.

*What is the impact of imputing missing data on the estimates of the total daily number of steps?*

The number of steps both median and mean increased.

## Are there differences in activity patterns between weekdays and weekends?

Time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

1. Create a factor variable with the levels "weekend" and "weekday"
2. Create grouping by in interval
3. Summarize the mean of steps
4. Plot Panel by weekday or weekend
```{r}
tidyActivity <- mutate(tidyActivity, dayType = factor(ifelse((wday(date)==1|wday(date)==7),"weekend","weekday")))
tidyActivityGroup <- group_by(tidyActivity,dayType, interval)
meanStepsByInterval <- summarize(tidyActivityGroup, meanSteps = mean(steps))
qplot(interval, meanSteps, data = meanStepsByInterval, facets = dayType ~ ., geom = "line", ylab ="Number of Steps")
```



